Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations

### 2020-07-25 paper.Rnw
finished "cutting" of paper and wrote appendices (minus updated explanation scatterplot which is the next task)

sine_poi_exp <- sine_poi_lime_default %>% plot_features()

@@

\begin{figure}[!t]
\centering
\includegraphics[width=2.25in]{./figure-static/bullets.png}
\caption{(Top left) Traditionally rifled gun barrel. The grooves and lands alternate to give bullets a spin during the firing process, which create markings (striations) on a bullet when fired. (Top right) Image of a fired bullet. The vertical stripes along the lower half of the bullet show groove and land engraved areas. The land engraved areas contain the microscopic striations created when the bullet passed through the barrel of the gun. (Bottom) Close up of a land engraved area showing striations (vertical lines).}
\label{fig:bullet}

@@

  # Create the plot
  bullet_pcp <- bullet_train %>%
    select(all_of(bullet_features), samesource, rfscore) %>%
    bind_rows(bullet_test %>%
                select(all_of(bullet_features), samesource, rfscore),
              .id = "set") %>%
    mutate(set = fct_recode(set,
                            "Training Set" = "1",
                            "Testing Set" = "2"),
           samesource = fct_recode(factor(samesource),
                                  "Same Source = TRUE" = "TRUE",
                                  "Same Source = FALSE" = "FALSE")) %>%
    rename("CCF" = "ccf", "CMS" = "cms", "Matches" = "matches",
           "Mismatches" = "mismatches", "Non-CMS" = "non_cms",
           "Rough Correlation" = "rough_cor", "Distance Std Dev" = "sd_D",
           "Distance" = "D", "Sum Peaks" = "sum_peaks") %>%
    na.omit() %>%
    ggplot(aes(color = rfscore)) +
    geom_pcp(aes(vars = vars(bullet_features_ordered$feature)),
             alpha = 0.2) +
    facet_grid(set ~ samesource) +
    scale_color_gradient2(low = "grey50",
                          high = "darkorange",
                          midpoint = 0.5) +
    theme_bw(base_family = "Times", base_size = 10) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          strip.placement = "outside",
          strip.background = element_rect(color = "white",
                                          fill = "white")) +
    labs(x = "Features Ordered by Random Forest Variable Importance \n(highest to lowest from left to right)",
         y = "Scaled Feature Values",
         color = "Random \nForest \nProbability")

@@

@
\caption{Parallel coordinate plots of the \citet{hare:2016} random forest predictions from the training (top facet) and testing (bottom facet) data. The observations are separated by same source (right column) and different source (left column) The y-axis shows the standardized feature values, and the x-axis shows the features used to fit the random forest (ordered by random forest impurity based feature importance). Each line corresponds to a comparison, and the line color represents the associated random forest probability. There are clear relationships between the feature values and the random forest probabilities.}
\label{fig:bullet-plot-rfpcp}
\end{figure*}

@@

To get an overview of the LIME explanations from the 36 applications, we consider a feature heatmap (\autoref{fig:bullet-feature-heatmap}). The plot is faceted vertically by simulation method and the Gower power and horizontally by the order the feature was chosen and whether the observation is a known same source or different source comparison. \kgc{This is where I left off...} This plot highlights several key features of the LIME explanations from the bullet matching dataset.

First of all, the density simulation methods produced the same explanations for almost all cases and LIME tuning parameters. As a result, the LIME explanations for the density methods are global and not local. Second, within a bin based simulation method, the features selected by LIME for an observation often vary based on the number of bins used but do not appear to vary based on the Gower power used. Especially with the equal bin explanations, there are vertical stripes that suggest a dependence of the LIME explanations on the number of bins used. The vertical stripes are not as apparent with the quantile bins. Lastly, there are clear differences between the LIME explanations for the cases that are matches and those that are not matches which are produced by the bin based simulation methods. This provides evidence that suggests that the features which the random forests uses to classify a match are different from the features used by the random forest to classify a non-match. The most important insight \autoref{fig:bullet-feature-heatmap} highlights is the dependence of the explanations on the LIME tuning parameter for many of the observations in the test data.

\begin{figure*}[!thp]
<<bullet-feature-heatmap, out.width = '\\textwidth', fig.width = 18, fig.height = 10.5>>=

# Create a feature heatmap
plot_feature_heatmap(bullet_explain %>%
                  mutate(label = as.factor(label)),
                facet_var = na.omit(bullet_test$samesource),
                order_method = "PCA") +
  scale_fill_gretchenalbrecht(palette = "last_rays", discrete = TRUE) +
  scale_color_gretchenalbrecht(palette = "last_rays", discrete = TRUE, reverse = TRUE) +
  theme_bw(base_family = "Times", base_size = 16) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.background = element_rect(color = "white", fill = "white"),
        strip.text.y.right = element_text(angle = 0), 
        legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = guide_legend(nrow = 1)) + 
  labs(y = "Case", color = "Complex Model Feature", fill = "Complex Model Feature")

@
\caption{Feature heatmap of the 36 LIME applications to the bullet comparison data test set. In addition to faceting the results by simulation method and order a feature was selected by LIME as done in \autoref{fig:sine-feature-heatmap}, this plot has been faceted by the Gower power and whether the the observation is from a match (TRUE) or non-match (FALSE). This plot shows several key insights about the LIME explanations. First, the features selected by LIME for the density simulation methods are mostly the same for all tuning parameters and all cases. Second, the features selected by LIME for the bin based methods are affected by the number of bins but not the Gower power. Third, the explanations produced by LIME using bin based methods result in different important features for the matches and non-matches.}
\label{fig:bullet-feature-heatmap}

@@

# Create a metric comparison plot
plot_metrics(bullet_explain %>% 
               mutate(label = as.factor(label)), 
             add_lines = TRUE) +
  theme_bw(base_family = "Times") +
  theme(strip.background = element_rect(color = "white", fill = "white"))
@
\caption{Plot of LIME assessment metrics for the 36 different applications of LIME applied to the bullet comparison data test set. The shape of the points represents the power used for the Gower distance calculation, and the color of the points represents the rank associated with that observation within a metric. The density simulation methods perform well for all metrics, but the bin based metrics often do not agree in terms of performance across metrics.}
\label{fig:bullet-assess-metric}
\end{figure*}

To try to identify the set of LIME tuning parameters with the most trustworthy set of explanations, we created an assessment metric plot shown in \autoref{fig:bullet-assess-metric}. The three metrics of average fidelity, average $R^2$, and MSEE were computed for each of the simulation methods and Gower power used to implement LIME. The plot is faceted by the simulation method, and the shape of a point represents the Gower power. The color of a point represents the rank within a metric. That is, the darker a point, the better performance of that method based on the metric.

The density simulation methods performed well across all tuning parameters. This is an interesting result, since \autoref{fig:bullet-feature-heatmap} shows that the density methods results in global explanations. The metric values for the bin based methods do not agree across metrics. For example, 2 and 3 quantile bins performed well according to the average $R^2$ but poorly according to MSEE. As a result, with the bin based methods, it is difficult to know which method to recommend. When considering the metrics across the Gower power used, the applications using a power of 0.5 perform best or as well as the other powers in across all simulation methods. This suggests that the power that leads to a more global explanation is preferred by LIME in this example.

Unfortunately, \autoref{fig:bullet-feature-heatmap} leaves us without a clear indication of which set of explanations to use to explain the random forest. Perhaps we can say that the kernel density method, 3 equal bins, and 3 quantile bins are some of the best performing methods. To better understand the explanations provided by these three methods, we take a closer look at explanations from these methods for two observations of interest where one is a known non-match (case \Sexpr{bullet_poi_nonmatch}) and the other is a known match (case \Sexpr{bullet_poi_match}).

Figures \ref{fig:bullet-eois-nonmatch} and \ref{fig:bullet-eois-match} contain plots that provide this closer look for the known non-match and match, respectively. The top row of plots in each figure are the explanation plots from \emph{lime}. The bottom row of plots show scatterplots of the features selected by LIME colored by the random forest model prediction. The location of the prediction of interest is indicated on the scatterplots by a diamond, and black solid lines are included on the plots representing the bin divisions for applications using bins. The bottom row plots are created using \emph{limeaid}. The plots within a column are associated with one of the three simulation methods (kernel density, 3 equal bins, or 3 quantile bins).

The scatterplots provide a visualization that helps to explain the LIME explanation. For example, consider the explanation for case \Sexpr{bullet_poi_nonmatch} when 3 quantile bins were used. The explanation plot from \emph{lime} shows that case \Sexpr{bullet_poi_nonmatch} having an observed CCF value greater than 0.320 supports a random forest prediction in favor of a match. The scatter plots show that many of the simulated values with CCF greater then 0.320 have random forest predictions greater than 0.5. Additionally, the LIME explanation indicates that a value of matches less than or equal to 1.66 contradicts a prediction in favor of a match. The scatter plots shows that almost all simulated values with a value of matches less than or equal to 1.66 have random forest predictions close to 0. Similar statements can be made about the 3 equal bin explanation for this case and the bin based methods for the known match observation explanations.

While the scatterplots provide visual explanations for the LIME explanations, some of them indicate that LIME falls short of a good explanation of the random forest predictions. Again, consider the explanation for case 325 when 3 quantile bins were used. From looking at the scatter plot of CCF versus matches, a better explanation for the random forest prediction would be to say that because the prediction of interest has a value of CCF less than 0.8 (approximately) and a value of matches less than 7 (approximately), the random forest is providing a prediction that supports a non-match. The relationship between ccf and rough correlation does not provide much evidence to support the random forest prediction one way or the other due to the mixture of random forest predictions ranging from 0 to 1 in most regressions, but there is a pentagon shaped region at the bottom of the scatter plot of matches versus rough correlation that the prediction of interest falls in that supports the random forest prediction of a non-match.

@@

eoi1_3qb <-
  plot_explain_scatter(bullet_explain_perms[1:3, ], alpha = 0.9, weights = TRUE) +
  scale_color_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  scale_fill_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  theme_bw(base_family = "Times", base_size = 8) +
  theme(
    strip.placement = "outside",
    strip.background = element_rect(color = "white",
                                    fill = "white")
  ) + 
  guides(linetype = guide_legend(override.aes = list(color = c("darkorange"))))

eoi2_3qb <-
  plot_explain_scatter(bullet_explain_perms[4:6, ], alpha = 0.9, weights = TRUE) +
  scale_color_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  scale_fill_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  theme_bw(base_family = "Times", base_size = 8) +
  theme(
    strip.placement = "outside",
    strip.background = element_rect(color = "white",
                                    fill = "white")
  ) + 
  guides(linetype = guide_legend(override.aes = list(color = c("grey50", "darkorange"))))

eoi1_3eb <-
  plot_explain_scatter(bullet_explain_perms[7:9, ], alpha = 0.9, weights = TRUE) +
  scale_color_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  scale_fill_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  theme_bw(base_family = "Times", base_size = 8) +
  theme(
    strip.placement = "outside",
    strip.background = element_rect(color = "white",
                                    fill = "white")
  ) + 
  guides(linetype = guide_legend(override.aes = list(color = c("grey50", "darkorange"))))

eoi2_3eb <-
  plot_explain_scatter(bullet_explain_perms[10:12, ], alpha = 0.9, weights = TRUE) +
  scale_color_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  scale_fill_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  theme_bw(base_family = "Times", base_size = 8) +
  theme(
    strip.placement = "outside",
    strip.background = element_rect(color = "white",
                                    fill = "white")
  ) + 
  guides(linetype = guide_legend(override.aes = list(color = c("grey50", "darkorange"))))

eoi1_kd <-
  plot_explain_scatter(bullet_explain_perms[13:15, ], alpha = 0.9, weights = TRUE) +
  scale_color_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  scale_fill_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  theme_bw(base_family = "Times", base_size = 8) +
  theme(
    strip.placement = "outside",
    strip.background = element_rect(color = "white",
                                    fill = "white")
  )

eoi2_kd <-
  plot_explain_scatter(bullet_explain_perms[16:18, ], alpha = 0.9, weights = TRUE) +
  scale_color_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  scale_fill_gradient2(
    low = "grey45",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  theme_bw(base_family = "Times", base_size = 8) +
  theme(
    strip.placement = "outside",
    strip.background = element_rect(color = "white",
                                    fill = "white")
  )

pf1_3qb <- plot_features(bullet_explain_perms[1:3,]) +
  scale_fill_manual(values = c("darkorange", "grey50"))
pf2_3qb <- plot_features(bullet_explain_perms[4:6,]) +
  scale_fill_manual(values = c("darkorange", "grey50"))
pf1_3eb <- plot_features(bullet_explain_perms[7:9,]) +
  scale_fill_manual(values = c("darkorange", "grey50"))
pf2_3eb <- plot_features(bullet_explain_perms[10:12,]) +
  scale_fill_manual(values = c("darkorange", "grey50"))
pf1_kd <- plot_features(bullet_explain_perms[13:15,]) +
  scale_fill_manual(values = c("darkorange", "grey50"))
pf2_kd <- plot_features(bullet_explain_perms[16:18,]) +
  scale_fill_manual(values = c("darkorange", "grey50"))
@

\begin{figure*}[!thp]
<<bullet-eois-nonmatch, out.width = '\\textwidth', fig.width = 15, fig.height = 7.8125, message = FALSE>>=
# Join the plots
plot_grid(pf2_3qb, pf2_3eb, pf2_kd,
          eoi2_3qb, eoi2_3eb, eoi2_kd,
          nrow = 2,
          rel_heights = c(0.45, 0.55))
@
\caption{Plots of LIME explanations for one case in the test data that is a known non-match. Each column contains plots associated with a different application (3 quantile bins, 3 equal bins, or kernel density). The top row of plots are the explanation plots from \emph{lime}. The bottom row shows scatter plots of the simulated data associated with the prediction of interest. The features plotted are those chosen by LIME in the explanations, and the points are colored by the the random forest predictions. Lines showing the divisions created in the feature space by the bin based LIME methods are included as solid black lines. The scatter plots can be used to better understand the LIME explanation and assess if it is a good explanation.}
\label{fig:bullet-eois-nonmatch}

<<bullet-eois-match, out.width = '\\textwidth', fig.width = 15, fig.height = 7.8125, message = FALSE>>=
# Join the plots
plot_grid(pf1_3qb, pf1_3eb, pf1_kd,
          eoi1_3qb, eoi1_3eb, eoi1_kd,
          nrow = 2,
          rel_heights = c(0.45, 0.55))
@
\caption{Plots with the same structure as \ref{fig:bullet-eois-match} but for an observation from the test data that is a known match.}
\label{fig:bullet-eois-match}
\end{figure*}

The scatter plots of the kernel density applications for both the match and non-match case show that almost all of the simulated values have random forest predictions less than 0.5, which provide evidence in favor of a non-match. It appears that simulation method is not providing enough values to cover the feature space that hit regions where the random forest provides predictions above 0.5. The training data has a much smaller amount of matches (\Sexpr{sum(bullet_train$samesource)}) than non-matches (\Sexpr{sum(bullet_train$samesource == FALSE)}), and this simulation method is clearly affected by this imbalance in the two classification categories. Furthermore, this may be the cause of the minimal amount of variability in the LIME explanations across all of the cases in the test data.

Without applying multiple applications LIME to the bullet test data or viewing diagnostic plots of the LIME explanations, it may be very possible to formulate reasons why the LIME explanations make sense. However, the sequence of plots in this section (Figures \ref{fig:bullet-feature-heatmap}, \ref{fig:bullet-lime-expls}, \ref{fig:bullet-assess-metric}, \ref{fig:bullet-eois-match}, and \ref{fig:bullet-eois-nonmatch}) suggest that we should be cautious to trust any of these LIME explanations. While some of them provide an explanation that may be reasonable, there could be better explanations (as seen in the 3 quantile bin case of \autoref{fig:bullet-eois-match}), and some of them provide explanations that make no sense (as seen in \autoref{fig:bullet-lime-expls}). It appears that either LIME needs to be further tuned to provide trustworthy and good explanations, or a different approach may provide better insight. As glimpsed with the scatter plots in Figures \ref{fig:bullet-eois-match} and \ref{fig:bullet-eois-nonmatch}, the more traditional approach to explanation random forest predictions by plotting the random forest predictions over the feature space (partial dependence plots), could provide better insight into the complex model with no need to assess the explanation method if the sampling method is well chosen.

@@

This paper highlights that while an explainer model is meant to provide clarity, it actually adds another layer of complexity to predictive models by requiring yet another model that needs to be assessed. Without an assessment of the explainer model, LIME is a black-box procedure of its own. We suggest the use of visual diagnostics to counteract the black-box nature of LIME and provide three diagnostic plots. Here we discuss the limitations and possible extensions of the diagnostic plots and concerns about LIME raise by applications of the diagnostics plots in this paper.

\subsection{Diagnostic Plot Limitations and Possible Extensions} \label{limit-exten}

% We acknowledge that not all prediction problems will produce trends as easy to interpret as the visualizations we have shown in the section, but we still believe that visualizations of the procedure can provide insight into the quality of the LIME explanation. We hope these figures will provide a starting point (if not an ending point) for the assessment of a single LIME explanation. Perhaps the inability to interpret a plot is itself telling, or the act of creating extensions of these visualizations to work with a different scenarios will lead to an understanding of the trustworthiness of the explanation.

%\hh{In terms of limitations of the visualizations: first two features span the x and y axis, additional features can be used for faceting. four features is completely feasible, beyond that the visualization might not work as well. The number of bins should be in the single digits. A large number of small multiples \citep{Becker:1996gy} needs a strong signal in the data to be effective - otherwise the visualization might be affected by change blindness \citep{simonsChangeBlindness1997, review19}. }

%The \data \ provide a \kge{neat} \kge{very clean} example of how to use the proposed visualizations to assess a single LIME explanation \kge{with two features. Once the number of features is greater than 3, adjustments will need to be made to the plots. In this case, multiple plots will be required to visualize the majority of the steps in the LIME procedure. Reasonable extensions would be scatterplot matrices or the use of faceting. \hh{XXX this is too vague: when do you use scatterplot matrices, what is the advantage of faceting. Make your solutions more concrete by discussing them as well as their limitations. }

%\hh{Both large number of observations and large number of features make visual assessments harder:} \kgc{I've decided to not discuss a large amount of data, because that really isn't a problem due to the size of the simulated dataset being chosen by the user, and there is not a need to go above 5000. Whereas, it is reasonable to expect that a user may select more than two features.} \hh{That the size of the data is chosen by the user is actually a good point to make. }

%\hh{the number of features can also be controlled by the user - lime is interested in a small number of features, as earlier mentioned, using faceting or scatterplot matrices we are able to deal with high single digits, low double digits of features. The limiting factors are screen space and cognitive load. }

>>>>

sine_poi_exp <- sine_poi_lime_default %>% plot_features() + theme(text = element_text(family = "Times", size = 10))

@@

\begin{figure}[!t]
\centering
\includegraphics[width=2in]{./figure-static/bullets.png}
\caption{(Top left) Traditionally rifled gun barrel. The grooves and lands alternate to give bullets a spin during the firing process, which create markings (striations) on a bullet when fired. (Top right) Image of a fired bullet. The vertical stripes along the lower half of the bullet show groove and land engraved areas. The land engraved areas contain the microscopic striations created when the bullet passed through the barrel of the gun. (Bottom) Close up of a land engraved area showing striations (vertical lines).}
\label{fig:bullet}

@@

  # Create the plot
  bullet_pcp <- bullet_train %>%
    select(all_of(bullet_features), samesource, rfscore) %>%
    bind_rows(bullet_test %>%
                select(all_of(bullet_features), samesource, rfscore),
              .id = "set") %>%
    mutate(set = fct_recode(set,
                            "Training Set" = "1",
                            "Testing Set" = "2"),
           samesource = fct_recode(factor(samesource),
                                  "Match" = "TRUE",
                                  "Non-Match" = "FALSE")) %>%
    rename("CCF" = "ccf", "CMS" = "cms", "Matches" = "matches",
           "Mismatches" = "mismatches", "Non-CMS" = "non_cms",
           "Rough Correlation" = "rough_cor", "Distance Std Dev" = "sd_D",
           "Distance" = "D", "Sum Peaks" = "sum_peaks") %>%
    na.omit() %>%
    ggplot(aes(color = rfscore)) +
    geom_pcp(aes(vars = vars(bullet_features_ordered$feature)),
             alpha = 0.2) +
    facet_grid(set ~ samesource) +
    scale_color_gradient2(low = "grey50",
                          high = "darkorange",
                          midpoint = 0.5) +
    theme_bw(base_family = "Times", base_size = 10) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          strip.placement = "outside",
          strip.background = element_rect(color = "white",
                                          fill = "white")) +
    labs(x = "Features Ordered by Random Forest Variable Importance \n(highest to lowest from left to right)",
         y = "Scaled Feature Values",
         color = "Random \nForest \nProbability")

@@

@
\caption{Parallel coordinate plots of the \citet{hare:2016} random forest predictions from the training (top facet) and testing (bottom facet) data. The observations are separated by known matching (right column) and non-matching (left column) signatures. The y-axis shows the standardized feature values, and the x-axis shows the features used to fit the random forest (ordered by random forest impurity based feature importance). Each line corresponds to an observation, and the line color represents the associated random forest probability. There are clear relationships between the feature values and the random forest probabilities.}
\label{fig:bullet-plot-rfpcp}
\end{figure*}

@@

To get an overview of the LIME explanations from the 36 applications, we consider a feature heatmap (\autoref{fig:bullet-feature-heatmap}). In addition to facets for simulation method and order of feature selected by LIME, this plot includes a vertical facet for Gower power and a horizontal facet for whether the observation is a known match or non-match. This plot highlights several key features of the LIME explanations from the bullet matching dataset.

First, the density simulation methods produce the same explanations for almost all cases and LIME tuning parameters suggesting the LIME explanations are global and not local. Second, within a bin based simulation method, the features selected by LIME for an observation often vary by the number of bins but do not appear to vary by the Gower power. With the equal bins, there are vertical stripes that suggest a dependence of the LIME explanations on the number of bins. The vertical stripes are not as apparent with the quantile bins. Lastly, there are clear differences between the LIME explanations produced by the bin based simulation methods for the matches and non-matches. This suggests that the features the random forest uses to classify a match or non-match are different.

\begin{figure*}[!thp]
<<bullet-feature-heatmap, out.width = '\\textwidth', fig.width = 18, fig.height = 12>>=
# Create a feature heatmap
plot_feature_heatmap(
  bullet_explain %>%
    mutate(
      label = as.factor(label),
      feature = fct_recode(
        feature,
        "Rough Correlation" = "rough_cor",
        "Consecutively Matching Striae" = "cms",
        "Distance" = "D",
        "Matches" = "matches",
        "Mismatches" = "mismatches",
        "Non-Consecutively Matching Striae" = "non_cms",
        "Cross Correlation Function" = "ccf",
        "Sum of Peaks" = "sum_peaks",
        "Distance Standard Deviation" = "sd_D"
      )
    ),
  facet_var = bullet_test %>%
    mutate(samesource = fct_recode(
      factor(samesource),
      "Match" = "TRUE",
      "Non-Match" = "FALSE"
    )) %>%
    pull(samesource) %>%
    na.omit(),
  order_method = "PCA"
) +
  scale_fill_gretchenalbrecht(palette = "last_rays", discrete = TRUE) +
  scale_color_gretchenalbrecht(palette = "last_rays",
                               discrete = TRUE,
                               reverse = TRUE) +
  theme_bw(base_family = "Times", base_size = 16) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.y.right = element_text(angle = 0),
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  guides(fill = guide_legend(nrow = 3)) +
  labs(y = "Case", color = "Complex Model Feature", fill = "Complex Model Feature")

@
\caption{Feature heatmap of the 36 LIME applications to the bullet comparison data test set. In addition to faceting the results by simulation method and LIME feature selection order, facets for the Gower power and whether the observation is a match or non-match are included. The vertical stripes of features selected indicate a dependence between the LIME explanations and tuning parameters.}
\label{fig:bullet-feature-heatmap}

@@

# Create a metric comparison plot
plot_metrics(bullet_explain %>% 
               mutate(label = as.factor(label)), 
             add_lines = TRUE) +
  theme_bw(base_family = "Times") +
  theme(strip.background = element_rect(color = "white", fill = "white"))
@
\caption{Plot of LIME assessment metrics for the applications of LIME on the bullet comparison data test set. The density simulation methods perform well for all metrics. The bin based metrics often do not agree in terms of performance across metrics.}
\label{fig:bullet-assess-metric}
\end{figure*}

To try to identify a set of LIME tuning parameters with the most trustworthy set of explanations, an assessment metric plot is considered (\autoref{fig:bullet-assess-metric}). All density simulation methods performed well based on the three metrics. This is an interesting result, since \autoref{fig:bullet-feature-heatmap} shows that the density methods results in global explanations. The metric values for the bin based methods do not agree across metrics. For example, 2 and 3 quantile bins performed well according to the average $R^2$ but poorly according to MSEE. As a result, with the bin based methods, it is difficult to know which method to recommend. The applications using a power of 0.5 perform the best or as well as the other powers across all simulation methods suggesting that the power that leads to a more global explanation is preferred by LIME.

Perhaps we can say that the kernel density method, 3 equal bins, or 3 quantile bins are some of the better performing methods, but unfortunately, \autoref{fig:bullet-assess-metric} leaves us without a clear indication of which set of explanations to use to explain the random forest. To better understand the explanations provided by these three methods, we take a closer look at explanations from these methods for two observations of interest, which will be referred to as case M (a known match) and case NM (a known non-match).

Figure \ref{fig:bullet-eois} include the explanation plots (from \emph{lime}) and explanation scatterplots for cases M and NM. Plots for 3 equal bins and 3 quantile bins are included. The scatterplots provide several insights into the LIME explanations. Here we focus on the explanation for case NM with 3 quantile bins, but similar statements can be made about the other plots.

The explanation plot from \emph{lime} for case NM with 3 quantile bins shows that an observed CCF value greater than 0.320 supports a random forest prediction in favor of a match, and the scatter plots show that many of the simulated values with CCF greater then 0.320 have random forest predictions greater than 0.5. Additionally, the LIME explanation indicates that a value of matches less than or equal to 1.66 contradicts a prediction in favor of a match, and the scatter plots show that almost all simulated values with a value of matches less than or equal to 1.66 have random forest predictions close to 0. While the LIME explanation makes sense based on the observations made from the explanation scatter plot, the plot also indicates that LIME falls short of providing a good explanation of the random forest prediction for case NM. Based on the scatter plot of CCF versus matches, a better explanation would be that because case NM has a value of CCF less than 0.8 and a value of matches less than 7, the random forest provides a prediction supporting a non-match. The relationship between CCF and rough correlation does not provide much evidence to support the random forest prediction one way or the other, but there is a pentagon shaped region at the bottom of the scatter plot of matches versus rough correlation with mostly predictions close to 0 that case NM falls and supports the random forest prediction of a non-match.

@@

eoi1_3qb <-
  plot_explain_scatter(bullet_explain_perms[1:3, ], alpha = 0.9, weights = TRUE) +
  scale_color_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  scale_fill_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  theme_bw(base_family = "Times", base_size = 10) +
  theme(
    strip.placement = "outside",
    strip.background = element_rect(color = "white",
                                    fill = "white")
  ) + 
  guides(linetype = guide_legend(override.aes = list(color = c("darkorange"))))

eoi2_3qb <-
  plot_explain_scatter(bullet_explain_perms[4:6, ], alpha = 0.9, weights = TRUE) +
  scale_color_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  scale_fill_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  theme_bw(base_family = "Times", base_size = 10) +
  theme(
    strip.placement = "outside",
    strip.background = element_rect(color = "white",
                                    fill = "white")
  ) + 
  guides(linetype = guide_legend(override.aes = list(color = c("grey50", "darkorange"))))

eoi1_3eb <-
  plot_explain_scatter(bullet_explain_perms[7:9, ], alpha = 0.9, weights = TRUE) +
  scale_color_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  scale_fill_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  theme_bw(base_family = "Times", base_size = 10) +
  theme(
    strip.placement = "outside",
    strip.background = element_rect(color = "white",
                                    fill = "white")
  ) + 
  guides(linetype = guide_legend(override.aes = list(color = c("grey50", "darkorange"))))

eoi2_3eb <-
  plot_explain_scatter(bullet_explain_perms[10:12, ], alpha = 0.9, weights = TRUE) +
  scale_color_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  scale_fill_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  theme_bw(base_family = "Times", base_size = 10) +
  theme(
    strip.placement = "outside",
    strip.background = element_rect(color = "white",
                                    fill = "white")
  ) + 
  guides(linetype = guide_legend(override.aes = list(color = c("grey50", "darkorange"))))

# eoi1_kd <-
#   plot_explain_scatter(bullet_explain_perms[13:15, ], alpha = 0.9, weights = TRUE) +
#   scale_color_gradient2(
#     low = "grey50",
#     high = "darkorange",
#     midpoint = 0.5,
#     limits = c(0, 1)
#   ) +
#   scale_fill_gradient2(
#     low = "grey50",
#     high = "darkorange",
#     midpoint = 0.5,
#     limits = c(0, 1)
#   ) +
#   theme_bw(base_family = "Times", base_size = 8) +
#   theme(
#     strip.placement = "outside",
#     strip.background = element_rect(color = "white",
#                                     fill = "white")
#   )
# 
# eoi2_kd <-
#   plot_explain_scatter(bullet_explain_perms[16:18, ], alpha = 0.9, weights = TRUE) +
#   scale_color_gradient2(
#     low = "grey50",
#     high = "darkorange",
#     midpoint = 0.5,
#     limits = c(0, 1)
#   ) +
#   scale_fill_gradient2(
#     low = "grey45",
#     high = "darkorange",
#     midpoint = 0.5,
#     limits = c(0, 1)
#   ) +
#   theme_bw(base_family = "Times", base_size = 8) +
#   theme(
#     strip.placement = "outside",
#     strip.background = element_rect(color = "white",
#                                     fill = "white")
#   )

pf1_3qb <- plot_features(bullet_explain_perms[1:3,]) +
  scale_fill_manual(values = c("darkorange", "grey50")) +
  theme(text = element_text(family = "Times", size = 10))
pf2_3qb <- plot_features(bullet_explain_perms[4:6,]) +
  scale_fill_manual(values = c("darkorange", "grey50")) + 
  theme(text = element_text(family = "Times", size = 10))
pf1_3eb <- plot_features(bullet_explain_perms[7:9,]) +
  scale_fill_manual(values = c("darkorange", "grey50")) +
  theme(text = element_text(family = "Times", size = 10))
pf2_3eb <- plot_features(bullet_explain_perms[10:12,]) +
  scale_fill_manual(values = c("darkorange", "grey50")) +
  theme(text = element_text(family = "Times", size = 10))
# pf1_kd <- plot_features(bullet_explain_perms[13:15,]) +
#   scale_fill_manual(values = c("darkorange", "grey50"))
# pf2_kd <- plot_features(bullet_explain_perms[16:18,]) +
#   scale_fill_manual(values = c("darkorange", "grey50"))
@

\begin{figure*}[!thp]
<<bullet-eois-nonmatch, out.width = '\\textwidth', fig.width = 12, fig.height = 17, fig.align = "center", out.width="90%", message = FALSE>>=
# Join the plots
plot_grid(
  pf2_3qb,
  pf2_3eb,
  eoi2_3qb,
  eoi2_3eb,
  pf1_3qb,
  pf1_3eb,
  eoi1_3qb,
  eoi1_3eb,
  nrow = 4,
  rel_heights = c(0.2, 0.3, 0.2, 0.3)
)
@
\caption{Plots of LIME explanations and explanation scatterplots for one case M and NM in the bullet test data. The first column contains plots associated with 3 quantile bins and the second with 3 equally spaced bins. The plots provide insights into the LIME explanations and allow for the assessment of the explanation quality.}
\label{fig:bullet-eois}
\end{figure*}

Without applying LIME with multiple tuning parameters to the bullet test data or viewing diagnostic plots of the LIME explanations, it may be very possible to formulate reasons why the LIME explanations make sense. However, the sequence of plots in this section (Figures \ref{fig:bullet-feature-heatmap}, \ref{fig:bullet-assess-metric}, and \ref{fig:bullet-eois}) suggest that we should be cautious to trust any of these LIME explanations. It appears that either LIME needs to be further tuned to provide trustworthy and good explanations, or a different approach may provide better insight.

@@

%\hh{The discussion should not have subsection ... but we can keep them for now for structural purposes.} \kg{Okay. I took them out.}

This paper highlights that while an explainer model is meant to provide clarity, it actually adds another layer of complexity to predictive models by requiring yet another model that needs to be assessed. Without an assessment of the explainer model, LIME is a black-box procedure of its own requiring blind trust in the explainer model. %\hh{XXX what is the implication? bring it to a punch: asking the user to trust into explainer...}
We suggest the use of visual diagnostics to counteract the black-box nature of LIME and provide three diagnostic plots.

The visualizations are intended to provide insight on how LIME works, assess the ability of the explainer model to capture the complex predictive model, and compare LIME explanations produced by different tuning parameters. While the visualizations accomplish these tasks, they also expose examples of the failings of LIME. To address the discovered failings of LIME, we will reconsider each of the claims about the performance of LIME made by \citet{ribeiro:2016} in light of the insights gained from the diagnostic visualizations.
